\input{Preamble_CV.tex}
\MyName{Jie Li}

\sepspace

%%% Personal details--------------------------------------------------------------------------

\SimpleEntry{\textbf{Github}: \href{https://github.com/jieli-matrix}{https://github.com/jieli-matrix}}
\SimpleEntry{\textbf{Website}: \href{https://jieli-matrix.github.io/}{https://jieli-matrix.github.io/}}
\SimpleEntry{\textbf{Email}: li\_j20@fudan.edu.cn}

%%% Education---------------------------------------------------------------------------------
\NewPart{Education}

\TimeEntryFocus{9.2020 - 6.2023}
{Master of Applied Mathematics }{at Fudan University}
{Supervised by Young PI \href{https://scholar.google.com/citations?user=bLERs80AAAAJ&hl=zh-CN}{Weiyang Ding}\\
Focus on Numerical Optimization and Matrix Computation}
\sepspace

\TimeEntry{9.2016 - 7.2020}
{Bachelor of Mathematics and Applied Mathematics}{at Lanzhou University}
\sepspace


%%% Skills and Qualifications --------------------------------------------------------------
\NewPart{Skills and Qualifications}
\NewSubPart{Programming Languages}
\TitleEntry{Advanced skills}{Python, Julia, Matlab}
\TitleEntry{Basic skills}{Git, Linux, Cpp}

\NewSubPart{Languages}
\TitleEntry{Native}{Chinese}
\TitleEntry{Advanced}{English (TOEFL:94 CET-4:632 CET-6:571)}

%%% Work experience------------------------------------------------------------------------
\NewPart{internship experience}
\TimeEntryDesc{10.2019 - 12.2019}{NLP Researcher in Core Development Platform}{at \href{https://www.iflytek.com/}{iFLYTEK}}
{\begin{compactitem}
    \item Automatic Pipeline on Senior High School Math Homework 
    \item Code Implementation of  \href{https://arxiv.org/abs/1911.06557}{Multi-Label Learning with Deep Forest}
\end{compactitem}
}



%%% Skills and Qualifications --------------------------------------------------------------
\NewPart{Projects}
\TimeEntryDesc{5.2021 - now}{Lowranksvd.jl}{ \href{https://github.com/jieli-matrix/Lowranksvd.jl/tree/dev}{Lowranksvd.jl}}
{\begin{compactitem}
    \item Algorithm 4.4 from \href{https://arxiv.org/abs/0909.4061}{Halko et al.} implemented by Julia with all CI tests passing.
    \item Algorithm 5.1 from \href{https://arxiv.org/abs/0909.4061}{Halko et al.} implemented by Julia with all CI tests passing.
\end{compactitem}
}

\TimeEntryDesc{4.2020}{CPUPredict}{\href{https://github.com/VealM/CPUPredict}{CPUPredict}}
{
    Time series prediction of the CPU's usage rates.  

    This project won the third place in Elastic Cloud College Challenge held by ALibaba.
}

\NewPart{Contests}
\TimeEntryDesc{11.2020}{Brain-Inspired Intelligence Contests held by Fudan University}{\quad}{
    \textbf{Efficent information encoding and energy efficiency in E/I balanced neuronal networks}  

    Numerical Simulation for E/I balanced neuronal networks.  
      
    I won the third place in the contest. 
}